A title

### 2022-01-02 pkg/oldgpuRandom/inst/doc/L2021_11/paper2.tex
paper2
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3082 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{article}

@@

<<preliminaries, echo=FALSE, include=TRUE>>=
options(continue="+  ",  prompt="R> ", digits=3, width=66)

opts_chunk$set(fig.height=3,
		fig.width=4.5,dev='png',tidy=TRUE,fig.align='center',background='#FFFFFF00', prompt=TRUE, highlight=FALSE)

@

@@

\author{
  Ruoyong Xu\\
  Department of Statistical Sciences\\
  University of Toronto\\
  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
  \texttt{ruoyong.xu@mail.utoronto.ca} \\
  %% examples of more authors
   \And
  Patrick Brown\\
  Department of Statistical Sciences\\
  University of Toronto\\
  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
  \texttt{patrick.brown@utoronto.ca} \\
}

\begin{document}
\maketitle
\setstretch{1.1}

@@

<<packages, results='hide', message=FALSE>>=
library(gpuRandom)
library(gpuR)
library(data.table)
library('geostatsp')
@

% swissrain matplot over betas
<<params, echo=FALSE, results='hide',eval=TRUE>>=
data('swissRain')
swissRain$elevation = extract(swissAltitude, swissRain)

params = rbind(
  c(shape=1.826358, range=20000, variance = 1, nugget = 0, anisoRatio = 10, anisoAngleRadians = 0.08),  
  c(shape=4, range=3.861873e+04, variance = 1, nugget = 15, anisoRatio = 16, anisoAngleRadians = 0.7),  
  c(shape=3, range=90*1000, variance = 1, nugget = 1.369707e-01, anisoRatio = 25, anisoAngleRadians = 0.3),
  c(shape=2, range=250*1000, variance = 1, nugget = 20, anisoRatio = 8.09, anisoAngleRadians = 0.5), 
  c(shape=0.2, range=15*1000, variance = 1, nugget = 0.02, anisoRatio = 65, anisoAngleRadians = 6.5),
  c(shape=0.3, range=20*1000, variance = 1, nugget = 0.08, anisoRatio = 45, anisoAngleRadians = 12.5)
)
colnames(params) <- c('shape','range', 'variance', 'nugget', 'anisoRatio', 'anisoAngleRadians')
@
<<showparams, eval=TRUE>>=
params
@

<<compute, echo=FALSE, results='hide', eval=TRUE>>=
result5<-gpuRandom::likfitLgmCov2d(
  data=swissRain,
  formula=rain~ elevation, 
  coordinates=swissRain@coords,
  params=params, 
  boxcox = c(1,0,4.962398e-01),
  cilevel=0.9,
  type = "double",
  reml=FALSE, 
  NparamPerIter=400,
  Nglobal=c(128,128),
  Nlocal=c(16,16),
  NlocalCache=2800,
  verbose=FALSE)

Betas <- seq(-12, 12, len=40)*1e-03

m <- length(Betas)
makeSymm <- function(m) {
  m[upper.tri(m)] <- t(m)[upper.tri(m)]
  return(m)
}
a=2 # which beta_i?
Nobs=result5$Nobs  # number of observations.
Ndata=result5$Ndata
Nparam=result5$Nparam
Ncov=result5$Ncov
XVYXVX=result5$XVYXVX
ssqY <- as.matrix(result5$ssqY)
detVar <- as.vector(result5$detVar)
detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
jacobian <- as.vector(result5$jacobian)
jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
Ucov <- Ncov-1
XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
XVY <- XVYXVX[ , 1:Ndata]

for (i in 1:Nparam){
  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
}

selectedrows <- (seq_len(Nparam)-1) * Ncov + a
XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
XVY_deleted <- XVY[-selectedrows, ]
X_aVY <- XVY[selectedrows, ]
X_aVX_a <- XTVinvX[selectedrows, a]

partA = matrix(0, nrow=Nparam, ncol=Ndata)
partB = matrix(0, nrow=Nparam, ncol=Ndata)
partC = matrix(0, nrow=Nparam, ncol=Ndata)
partD = matrix(0, nrow=Nparam, ncol=Ndata)
partE = matrix(0, nrow=Nparam, ncol=Ndata)

for (i in 1:Nparam){
  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
  temp <- solve(XTVinvX_deleted[interval, ])

  # part (A) have 2 data sets
  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
  # part (B) have 2 data sets.   has beta
  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
  # part (C) no data sets.  has beta
  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
  # part (D) have 2 data sets.    has beta
  partD[i,] = -2* X_aVY[i, ]
  # part (E)
  partE[i,] = X_aVX_a[i]
}

All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
for (bet in 1:m){
  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
}
colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5','6')
@

<<matplot_beta, eval=TRUE,echo=FALSE, fig.height =4.5, fig.width=6, cache=TRUE, fig.cap='profile log-Likelihoods of slope parameter for six covariance parameter sets',fig.pos='H'>>=
matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)
@
% swissrain matplot over betas

%meuse data
<<compare_mle, eval=TRUE,echo=FALSE>>=
rm(list=ls(all=TRUE))
library("sp")
data(meuse)
class(meuse)
coordinates(meuse) <- c("x", "y")
class(meuse)
library('geostatsp')
meuse$Y = log(meuse$zinc)
meuse$sqrtdist = sqrt(meuse$dist)

Res1 =  geostatsp::lgm( formula=Y~sqrtdist, 
                        data=meuse, 
                        grid=20,
                        reml = FALSE,
                        fixBoxcox=FALSE, fixShape=FALSE, fixNugget = FALSE,  #Set to FALSE to estimate the nugget effect parameter.
                        aniso=TRUE )
Res1$summary[,c('estimate','ci0.005', 'ci0.995')]
Res1$optim$mle
@

\pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.
<<compare_gpurandom, cache=FALSE, eval=FALSE,echo=FALSE>>=
library('gpuRandom')
library('gpuR')
  ParamList = list(
    range = c(exp(seq(log(100), log(173.1643928), len=6)), exp(seq(log(185), log(330), len=8))),
    shape = c(seq(0.1,3.5, len=4), seq(4, 6, len=4)),
    anisoRatio =c( seq(1, 2.301060332, len=3), seq(2.8, 7, len=4)),
    anisoAngleRadians = c( seq(0, 0.274896758, len=4), seq(0.4, 0.8, len=4)),
    nugget = c(seq(0,0.3946587, len=3), seq(0.5, 3, len=5))
  ) 
  params = do.call(expand.grid, ParamList)

result3<- gpuRandom::likfitLgmCov(spatialmodel=Res1,
                           params=params, # CPU matrix for now, users need to provide proper parameters given their specific need
                           boxcox=c(seq(-3, -0.903901505, len=4), seq(-0.5,1,len=4)),  
                           paramToEstimate = c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'), 
                           cilevel=0.95, 
                           type = "double",
                           reml=FALSE, 
                           NparamPerIter=400,
                           Nglobal=c(128,64),
                           Nlocal=c(16,16),
                           NlocalCache=2800,
                           verbose=FALSE)

cbind(Res1$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),c('estimate','ci0.005', 'ci0.995')], result3$summary[c('range','shape','nugget','anisoRatio','anisoAngleRadians','boxcox'),])
@

<<matplotmeuse, echo=FALSE, eval=TRUE>>=
result1<-gpuRandom::getProfLogL(data=meuse,
                                formula=Y~sqrtdist, 
                                coordinates=meuse@coords,
                                params=params, 
                                boxcox = -0.903901505,
                                type = "double",
                                NparamPerIter=400,
                                gpuElementsOnly=FALSE,
                                reml=FALSE, 
                                Nglobal=c(128,64),
                                Nlocal=c(16,16),
                                NlocalCache=2800)

LogLik = result1$LogLik[seq(1,50176,400),3]
length(LogLik) 
Betas <- seq(-12, 12, len=40)*1e-03
m <- length(Betas)
makeSymm <- function(m) {
  m[upper.tri(m)] <- t(m)[upper.tri(m)]
  return(m)
}
a=2 # which beta_i?
Nobs=result1$Nobs  # number of observations.
Ndata=result1$Ndata
Nparam=result1$Nparam[seq(1,50176,400),]
Ncov=result1$Ncov
XVYXVX=result1$XVYXVX
ssqY <- as.matrix(result1$ssqY)
detVar <- as.vector(result1$detVar)
detVar <- matrix(rep(detVar, Ndata), nrow=Nparam)
jacobian <- as.vector(result1$jacobian)
jacobian <- do.call(rbind, replicate(Nparam, jacobian, simplify = FALSE))  #dim(XVYXVX)
Ucov <- Ncov-1
XTVinvX <- XVYXVX[ , (ncol(XVYXVX)-Ncov+1):ncol(XVYXVX)]
XVY <- XVYXVX[ , 1:Ndata]

for (i in 1:Nparam){
  XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ] <- makeSymm(XTVinvX[((i-1)*Ncov+1) : (i*Ncov), ])
}

selectedrows <- (seq_len(Nparam)-1) * Ncov + a
XTVinvX_deleted <- matrix(XTVinvX[-selectedrows,-a], ncol=Ucov, byrow=TRUE)
XTVinvX_a <- matrix(XTVinvX[selectedrows, -a], nrow=Nparam, ncol=Ucov)
#XTVinvX_a <- matrix(XTVinvX[-selectedrows, a], nrow=Nparam*Ucov, ncol=1)[1:10,]
XVY_deleted <- XVY[-selectedrows, ]
X_aVY <- XVY[selectedrows, ]
X_aVX_a <- XTVinvX[selectedrows, a]

partA = matrix(0, nrow=Nparam, ncol=Ndata)
partB = matrix(0, nrow=Nparam, ncol=Ndata)
partC = matrix(0, nrow=Nparam, ncol=Ndata)
partD = matrix(0, nrow=Nparam, ncol=Ndata)
partE = matrix(0, nrow=Nparam, ncol=Ndata)

for (i in 1:Nparam){
  interval <- c(((i-1)*Ucov+1) : (i*Ucov))
  temp <- solve(XTVinvX_deleted[interval, ])

  # part (A) have 2 data sets
  partA[i,] = rowSums(XVY_deleted[interval,] %*% temp) * XVY_deleted[interval,]
  # part (B) have 2 data sets.   has beta
  partB[i,] = -2*XVY_deleted[interval,] %*% temp %*% XTVinvX_a[i,]
  # part (C) no data sets.  has beta
  partC[i,] = XTVinvX_a[i, ] %*% temp %*% XTVinvX_a[i, ]
  #partC[i,] = XTVinvX_a[interval, ] %*% temp %*% XTVinvX_a[interval, ]
  # part (D) have 2 data sets.    has beta
  partD[i,] = -2* X_aVY[i, ]
  # part (E)
  partE[i,] = X_aVX_a[i]
}

All_min2loglik_forthisbeta = matrix(0, nrow=m, ncol=Nparam)
for (bet in 1:m){
  ssqResidual <- ssqY + Betas[bet] *partD + Betas[bet]^2 *partE - (partA + Betas[bet]* partB + Betas[bet]^2 * partC)
    All_min2loglik_forthisbeta[bet,] <- -0.5*(Nobs*log(ssqResidual/Nobs) + detVar + Nobs + Nobs*log(2*pi) + jacobian)[,3]
}
colnames(All_min2loglik_forthisbeta) <- c('1','2','3','4','5', '6')
matplot(Betas,All_min2loglik_forthisbeta, type="l", lty=1, xlab="beta_2", ylab = 'profile LogL')
legend("topleft", cex = 0.75, legend = colnames(All_min2loglik_forthisbeta)[1:6], col = 1:6, lty = 1)

@
%meuse data

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\makeatletter
\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}}}
\makeatother
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\makeatletter
\@ifundefined{AddToHook}{}{\AddToHook{package/xcolor/after}{
\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
}}
\makeatother
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

@@

\author{
  Ruoyong Xu\\
  Department of Statistical Sciences\\
  University of Toronto\\
  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
  \texttt{ruoyong.xu@mail.utoronto.ca} \\
  %% examples of more authors
   \And
  Patrick Brown\\
  Department of Statistical Sciences\\
  University of Toronto\\
  700 University Ave., Toronto, ON M5G 1Z5, Canada\\
  \texttt{patrick.brown@utoronto.ca} \\
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
\setstretch{1.1}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> library(gpuRandom)
R> library(gpuR)
R> library(data.table)
R> library("geostatsp")
\end{verbatim}
\end{kframe}
\end{knitrout}

% swissrain matplot over betas

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
R> params
##      shape  range variance nugget anisoRatio anisoAngleRadians
## [1,]  1.83  20000        1  0.000      10.00              0.08
## [2,]  4.00  38619        1 15.000      16.00              0.70
## [3,]  3.00  90000        1  0.137      25.00              0.30
## [4,]  2.00 250000        1 20.000       8.09              0.50
## [5,]  0.20  15000        1  0.020      65.00              6.50
## [6,]  0.30  20000        1  0.080      45.00             12.50
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}[H]

{\centering \includegraphics[width=\maxwidth]{figure/matplot_beta-1}

}

\caption[profile log-Likelihoods of slope parameter for six covariance parameter sets]{profile log-Likelihoods of slope parameter for six covariance parameter sets}\label{fig:matplot_beta}
\end{figure}

\end{knitrout}
% swissrain matplot over betas

%meuse data
\begin{knitrout}
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## [1] "data.frame"
## [1] "SpatialPointsDataFrame"
## attr(,"package")
## [1] "sp"
##                    estimate   ci0.005 ci0.995
## (Intercept)         0.91634  0.121205   1.711
## sqrtdist           -0.08445 -1.623508   1.455
## sdNugget            0.00785  0.000379   0.163
## sdSpatial           0.01250  0.000690   0.227
## range             173.16439 92.775636 323.209
## shape               4.00000 -9.880685  17.881
## anisoRatio          2.30106  1.138810   4.649
## anisoAngleRadians   0.27490 -0.167582   0.717
## anisoAngleDegrees  15.75042 -9.601751  41.103
## boxcox             -0.90390 -2.558961   0.751
##            nugget          variance             range 
##             0.395             1.000           173.164 
##             shape        anisoRatio anisoAngleRadians 
##             4.000             2.301             0.275 
##            boxcox 
##            -0.904
\end{verbatim}
\end{kframe}
\end{knitrout}

\pkg{gpuBatchMatrix} give wider ci and more accurate MLE's, expecially for the shape parameter.

%meuse data
